---
title: "Homework2"
author: "Akhil Jose Chandy"
date: "1/23/2020"
output: word_document
---

```{r setup, include=FALSE}
library(rio) #For importing datsets   
library(stargazer)  #for nice outputs
library(dplyr) #useful data manipulation
library(ggplot2)  #good for plots
library(reshape2) #useful in converting tables from long to wide format and vice-versa
library(scales) # For rescaling the graphs
library(lattice) #for making conditional plots
library(reshape2)
library(plyr)
library(dplyr)
library(corrplot)
```

```{r}
rm(list = ls())  #hygiene practice-- clean up your worlspace in R

library(readxl)  #readxl is part of readxl package
setwd("C:/Akhil/USF BAIS/SDM/Working Directory")
dataset <- import("baldridge2011.xlsx")
colnames(dataset) <- tolower(colnames(dataset))  #forcing any capitalization in varnames to lowercase
library(car)
#brief(d)  #car function
# str(d)   #shows data structure
dataset$ccrtotal = as.numeric(dataset$ccrtotal)     #forcing var to be numeric
dataset$ccat4total = as.numeric(dataset$ccat4total)
dataset$icat4total = as.numeric(dataset$icat4total)
dataset$ccat7total = as.numeric(dataset$ccat7total)
#dataset$icat2total[is.numeric(dataset$icat2total)]
```
Question 1)
```{r}
m1_data=subset(dataset,year >= 1999 & year <= 2006 ,select=c(ccat7total,icat4total))
m1=lm(ccat7total ~ icat4total, data=m1_data)
stargazer(m1,type="text",
 dep.var.labels=c("ccat7total"),
 covariate.labels=c("icat4total"))
stargazer(type = "text",title="Correlation Matrix", cor(m1_data,use="complete.obs"))
summary(m1)
```
The model is statistically sginificant with a p value of 8.948e-08

A unit increase in icat4total is associated with an increase in the value of ccat7total by 2.44 in addition to the value of the intercept(70.75)

For a simple linear regression with only one independent the R squared value is the square of the value of correlation between the independent and dependent, in this case that is (0.335)^2,which is equal to 0.112

Question 2)
```{r}
m2_data=subset(dataset,year >= 1999 & year <= 2006 ,select=c(ccat7total,icat1total, icat2total, icat3total, icat4total, icat5total, icat6total))
m2=lm(ccat7total ~ icat1total+icat2total+icat3total+icat4total+icat5total+icat6total, data=m2_data)
#summary(m2)
stargazer(m2,type="text",
dep.var.labels=c("ccat7total"),
covariate.labels=c("icat1total", "icat2total","icat3total","icat4total","icat5total","icat6total"))
summary(m2)

```
The model is statistically significant.

The icat1total is statistically significant(at alpha = 0.05 ) with a p value close to 0.
Controling for the other variables the value of ccat7total is associated with an increases of 1.68 for an unit increase in icat1total.

The icat6total is statistically significant(at alpha = 0.10 ) with a p value 0.0679 .
Controling for the other variables the value of ccat7total is associated with an increases of 0.9533 for an unit increase in icat6total.

The icat3total is statistically significant(at alpha = 0.10 ) with a p value 0.0951 .
Controling for the other variables the value of ccat7total is associated with an increases of 0.9942 for an unit increase in icat3total

The model has an adjusted R squared of 0.27 which doesn't seem to be very good(although R squared alone cannot be used to judge the fit of the model)

Residual standard error is another goodness of fit measure and it has a value of 44.47 scores.
It means that the standard distance between the fitted line and the observed scores for ccat7total are 44.47, which in my opinion seems like a large difference.


Question 3)
```{r}
m3_data=subset(dataset,year >= 1999 & year <= 2006 ,select=c(ccat7total,year,sector,icat1total, icat2total, icat3total, icat4total, icat5total, icat6total))
# convert sector to words to make it readable
m3_data$sector=factor(m3_data$sector,labels=c("mfg","service","small","education",
                                        "health","nonprofit"))


m3 <- lm(ccat7total ~ icat1total+icat2total+icat3total+icat4total+icat5total+ icat6total+as.factor(year)+ as.factor(sector), data=m3_data)
#summary(m3)
stargazer(m3,type="text",
dep.var.labels=c("ccat7total"),
covariate.labels=c("icat1total","icat2total","icat3total","icat4total","icat5total","icat6total",
                  "2000","2001","2002","2003",   
                  "2004","2005","2006","service","small","education","health","non profit"))
summary(m3)
```
The icat1total is statistically significant(at alpha = 0.10 ) with a p value close to 0.0726 .
Controling for the other variables the value of ccat7total is associated with an increases of 0.9964 for an unit increase in icat1total.

The icat6total is statistically significant(at alpha = 0.05 ) with a p value 0.0271  .
Controling for the other variables the value of ccat7total is associated with an increases of 1.2455 for an unit increase in icat6total.

The year categories are not statistically significant.

However the base level of year 1999 is significant and is a part of the intercept.According to
my model ,for an organization in the manufacturing sector, in the year 1999 the ccat7total would be lower than for any other year by 75.4096, provided we control for all other variables in the model.

We see that the sectors of small and education are statistically significant and the intercept is also significant. The manufacturing sector is a part of the intercept.
Therefore, in the year 1999, for an organization in the manufacturing sector the ccat7total would be lower by 75.4096 than for an organization in another sector provided we control for all other variables.

We also see that in the year 1999, the ccat7total would be higher for an organization in the small sector by 28.6906 than for an organization in the manufacturing sector, controlling for all other variables.

We also see that in the year 1999, the ccat7total would be higher for an organization in the  education sector by 21.0497 than for an organization in the manufacturing sector, controlling for all other variables.


Question 4a)
```{r}
m4_data=subset(dataset,year >= 1999 & year <= 2006 ,select=c(ccat7total,year,sector,icat1total, icat2total, icat3total, icat4total, icat5total, icat6total))
#convert sector to words to make it readable
m4_data$sector=factor(m4_data$sector,labels=c("mfg","service","small","education",
                                        "health","nonprofit"))
m4=lm(ccat7total~icat1total+icat2total+icat3total+icat4total+icat5total+ icat6total+as.factor(year)+ as.factor(sector)+icat1total*icat6total, data=m4_data)
summary(m4)
```
a)
I would say model 3a is better due to parsimony. Further the addition of the interaction term does not significantly increase the R square value nor does it significantly reduce the residuals standar errors.
b)
Based on this model we could say that icat1total together with icat7total does have an impact on the value of ccat7total. But by themselves they do not seem to have an effect.
c)
The icat1total and icat7total do not seem to be significant even at the 10% mark.

From the model,
Controlling for all other variables (icat6total=0) the ccat7total would decrease by 
2.79 ,in the year 1999 for an organization in the manufacturing sector, for an unit increase in icat1total.

Controlling for all other variables (icat1total=0) the ccat7total would decrease by 
4.02 ,in the year 1999 for an organization in the manufacturing sector, for an unit increase in icat6total.

For non zero values of icat1toal and icat6total and controlling for all other variables, the value of ccat7total ,in the year 1999 for an organization in the manufacturing sector, would be lower by ((-2.79721)+(-4.02437) + 0.08019) 6.73 for an unit increase in value of icat1total and icat7total.



Question 4b)
```{r}
m4_data=subset(dataset,year >= 1999 & year <= 2006 ,select=c(ccat7total,year,sector,icat1total, icat2total, icat3total, icat4total, icat5total, icat6total))


#Identify median of icat1total and icat6total
stargazer(m4_data[c("icat1total","icat6total")],type = "text", 
          digits=2, median = TRUE)


m4b=lm(ccat7total~icat1total+icat2total+icat3total+icat4total+icat5total+ icat6total+as.factor(year)+ as.factor(sector)+I((icat1total-60)*(icat6total-43)), data=m4_data)
summary(m4b)

```
Controlling for all other variables the ccat7total would increase by 
0.65090 for an unit increase in icat1total,in the year 1999 for an organization in the manufacturing sector with a icat6total =43 

Controlling for all other variables the ccat7total would increase by 
0.78695 for an unit increase in icat6total,in the year 1999 for an organization in the manufacturing sector with an icat1total = 60

For  values of icat1toal = 60 and icat6total = 43 and controlling for all other variables, the value of ccat7total ,in the year 1999 for an organization in the manufacturing sector, would be higher by ((0.65090) +(0.78695)+ 0.08019) 1.52 for an unit increase in value of icat1total and icat7total.

Question 5)
```{r}
m5_data=subset(dataset,year >= 1999 & year <= 2006 ,select=c(ccat7total,year,sector,icat1total, icat2total, icat3total, icat4total, icat5total, icat6total))

#convert sector to words to make it readable
m5_data$sector=factor(m5_data$sector,labels=c("mfg","service","small","education",
                                        "health","nonprofit"))

#Creating a dummy for non profit
m5_data$nonprofit <- ifelse(m5_data$sector=="nonprofit", 1, 0)
m5_data$sector=as.factor(m5_data$sector)

m5=lm(ccat7total~icat1total+icat2total+icat3total+icat4total+icat5total+ icat6total+as.factor(year)+ as.factor(sector)+I(icat6total*(sector=="nonprofit")), data=m5_data)

summary(m5)

stargazer(m5,type="text",
dep.var.labels=c("ccat7total"),
covariate.labels=c("icat1total","icat2total","icat3total","icat4total","icat5total","icat6total",
                  "2000","2001","2002","2003",   
                  "2004","2005","2006","service","small","education","health","non profit","Interaction of icat6total&Binary Non profit"))
```
Interaction term is signnificant.

Interpreting the interaction term

Controlling for all other variables, an organization in the non profit sector with a non zero value of icat6toal, will have a ccat7total score that is different  by (-137.20 +3.012)  than an organization in the manufacturing sector in the year 1999 for the same value of icat6toal.

or
Controlling for all other variables,
for two companies in the non profit sector the value of ccat7total is expected to be (3.012 + 1.020 ) 4.032 higher than the company with one unit lower score of icat6total.

In comparison with m3

The magnitude of y intercept is significantly smaller for m5.
The coefficient for non profit sector is now significant and has a magnitude of 137.2035.


Question 6
```{r}
m6_data=subset(dataset,year >= 1999 & year <= 2006 ,select=c(ccat7total,year,sector,icat1total, icat2total, icat3total, icat4total, icat5total, icat6total))

#convert sector to words to make it readable
m6_data$sector=factor(m6_data$sector,labels=c("mfg","service","small","education",
                                        "health","nonprofit"))
#Creating a dummy for non profit
m6_data$year=as.factor(m6_data$year)

str(m6_data)

m6=lm(ccat7total~icat1total+icat2total+icat3total+icat4total+icat5total+ icat6total+as.factor(year)+ as.factor(sector)+I(icat6total*(year=="2006")), data=m6_data)

summary(m6)

stargazer(m6,type="text",
dep.var.labels=c("ccat7total"),
covariate.labels=c("icat1total","icat2total","icat3total","icat4total","icat5total","icat6total",
                  "2000","2001","2002","2003",   
                  "2004","2005","2006","service","small","education","health","non profit","Interaction of icat6total&Binary for year 2006"))


```
Interpreting interaction

Controlling for all other variables,
for two companies in the year 2006 the value of ccat7total is expected to be (2.57 + 0.953 ) 3.523 higher than the company with one unit lower score of icat6total.

In comparison with m3
The magnitude of y intercept is significantly smaller for m6.
The coefficient for non profit sector is now significant and has a magnitude of 122.3979.    


Question 7)
```{r}
a=data.frame("Model 1",summary(m1)$r.squared,summary(m1)$adj.r.squared,AIC(m1),BIC(m1))
names(a)=c("Model number","R squared","Adj. Rsquared","AIC","BIC")

b=data.frame("Model 2",summary(m2)$r.squared,summary(m2)$adj.r.squared,AIC(m2),BIC(m2))
names(b)=c("Model number","R squared","Adj. Rsquared","AIC","BIC")

c=data.frame("Model 3",summary(m3)$r.squared,summary(m3)$adj.r.squared,AIC(m3),BIC(m3))
names(c)=c("Model number","R squared","Adj. Rsquared","AIC","BIC")

d=data.frame("Model 4",summary(m4)$r.squared,summary(m4)$adj.r.squared,AIC(m4),BIC(m4))
names(d)=c("Model number","R squared","Adj. Rsquared","AIC","BIC")

d_b=data.frame("Model 4b",summary(m4b)$r.squared,summary(m4b)$adj.r.squared,AIC(m4b),BIC(m4b))
names(d_b)=c("Model number","R squared","Adj. Rsquared","AIC","BIC")

e=data.frame("Model 5",summary(m5)$r.squared,summary(m5)$adj.r.squared,AIC(m5),BIC(m5))
names(e)=c("Model number","R squared","Adj. Rsquared","AIC","BIC")

f=data.frame("Model 6",summary(m6)$r.squared,summary(m6)$adj.r.squared,AIC(m6),BIC(m6))
names(f)=c("Model number","R squared","Adj. Rsquared","AIC","BIC")

comparison=rbind(a,b,c,d,d_b,e,f)

stargazer(comparison,type="text",summary=FALSE,digits=2)

```

```{r}
stargazer(vif(m2),type="text",digits=2)
stargazer(vif(m3),vif(m4),vif(m4b),vif(m5),vif(m6),type="text",digits=2)
```
From the tables above we see that all models greater than model 4 have similar R square values
among these models models 4 and models 4a have the lower AIC and BIC values.

When we look at the VIF table model 4b has the lowest inflation of its coefficients due to multicollinearity.

Based on the above factors I'd choose model 4b as my best model.

Question 8)
```{r}
#Summary Stats
m2_summary=subset(dataset,year >= 1999 & year <= 2006 ,
                  select=c(ccat7total,icat1total, icat2total, 
                           icat3total, icat4total, icat5total, icat6total,sector))



m2_summary$sector=as.factor(m2_summary$sector)

dfm <- melt(m2_summary, id.vars = c("sector"),na.rm=TRUE)

dis=lapply(list(.(sector, variable)), 
       ddply, .data = dfm, .fun = summarise, 
       mean = mean(value), 
       sd = sd(value),
       N=length(value),
       se=sd/sqrt(N))


stargazer(dis,type="text",summary=FALSE,digits=2)


#Correlation

m2_summary = na.omit(m2_summary)
m2_summary <- m2_summary %>% mutate_if(is.character, as.numeric)


library(xtable)
# x is a matrix containing the data
# method : correlation method. "pearson"" or "spearman"" is supported
# removeTriangle : remove upper or lower triangle
# results :  if "html" or "latex"
  # the results will be displayed in html or latex format
corstars <-function(x, method=c("pearson", "spearman"), removeTriangle=c("upper", "lower"),
                     result=c("none", "html", "latex")){
    #Compute correlation matrix
    require(Hmisc)
    x <- as.matrix(x)
    correlation_matrix<-rcorr(x, type=method[1])
    R <- correlation_matrix$r # Matrix of correlation coeficients
    p <- correlation_matrix$P # Matrix of p-value 
    
    ## Define notions for significance levels; spacing is important.
    mystars <- ifelse(p < .0001, "****", ifelse(p < .001, "*** ", ifelse(p < .01, "**  ", ifelse(p < .05, "*   ", "    "))))
    
    ## trunctuate the correlation matrix to two decimal
    R <- format(round(cbind(rep(-1.11, ncol(x)), R), 2))[,-1]
    
    ## build a new matrix that includes the correlations with their apropriate stars
    Rnew <- matrix(paste(R, mystars, sep=""), ncol=ncol(x))
    diag(Rnew) <- paste(diag(R), " ", sep="")
    rownames(Rnew) <- colnames(x)
    colnames(Rnew) <- paste(colnames(x), "", sep="")
    
    ## remove upper triangle of correlation matrix
    if(removeTriangle[1]=="upper"){
      Rnew <- as.matrix(Rnew)
      Rnew[upper.tri(Rnew, diag = TRUE)] <- ""
      Rnew <- as.data.frame(Rnew)
    }
    
    ## remove lower triangle of correlation matrix
    else if(removeTriangle[1]=="lower"){
      Rnew <- as.matrix(Rnew)
      Rnew[lower.tri(Rnew, diag = TRUE)] <- ""
      Rnew <- as.data.frame(Rnew)
    }
    
    ## remove last column and return the correlation matrix
    Rnew <- cbind(Rnew[1:length(Rnew)-1])
    if (result[1]=="none") return(Rnew)
    else{
      if(result[1]=="html") print(xtable(Rnew), type="html")
      else print(xtable(Rnew), type="latex") 
    }
} 
corr=corstars(m2_summary, result="none")   # **** p < .0001,***p < .001,** p < .01, * p < .05 

stargazer(corr, type = "text", title="Correlations", 
                    summary=FALSE)
cat("\n**** p < .0001,***p < .001,** p < .01, * p < .05")

```
Question 9
```{r}

mbest_data=subset(dataset,year >= 1999 & year <= 2006 ,select=c(ccat7total,year,sector,icat1total, icat2total, icat3total, icat4total, icat5total, icat6total))

#Model with all variables
m_best1=lm(ccat7total~icat1total+icat2total+icat3total+icat4total+icat6total+as.factor(year)+ as.factor(sector),data=mbest_data)

summary(m_best1)

#Checking correlations to identify terms that have good correlations with each other and the ccat7total
mbest_data = na.omit(mbest_data)
mbest_data <- mbest_data %>% mutate_if(is.character, as.numeric)

corr2=corstars(mbest_data, result="none")   # **** p < .0001,***p < .001,** p < .01, * p < .05 

stargazer(corr2, type = "text", title="Correlations", 
                    summary=FALSE)

#Choose icat1total+icat2total+icat3total+icat6total
m_best2=lm(ccat7total~icat1total+icat2total+icat3total+icat6total+as.factor(year)+ as.factor(sector),data=mbest_data)


#summary(m_best2)

#Removing factor variables (Rsquared drops)

m_best3=lm(ccat7total~icat1total+icat2total+icat3total+icat6total,data=mbest_data)

#summary(m_best3)

#Removing icat2total variable and including year factor (Rsquared rises)

m_best5=lm(ccat7total~icat1total+icat3total+icat6total+as.factor(year),data=mbest_data)

#summary(m_best5)

# Finding the median for mean centering

stargazer(mbest_data[c("icat1total","icat6total")],type = "text", 
          digits=2, median = TRUE)


mbest_data$year=as.factor(mbest_data$year)
mbest_data$sector=as.factor(mbest_data$sector)

#Best model based on multiple trials

m_best6=lm(ccat7total~icat1total+icat3total+icat6total+year+as.factor(sector)+
             I((icat1total-60)*(icat6total-43))+I((icat6total-43)*(sector=="6")),data=mbest_data)

summary(m_best6)

# Acceptable VIF values
vif(m_best6)

d=as.data.frame(AIC(m_best1,m_best2,m_best3,m_best5,m_best6))

#Also see that the model 6 has the lowest AIC values
stargazer(d,type="text",summary=FALSE)
```

